---
title: "stringsAsFactors = HELLNO"
author: "Hesham Mahrous"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{stringsAsFactors = HELLNO}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

### Factors

Factors are a very useful type of variable in R, but they can also drive you nuts. Especially the "stealth factor" that you think of as character.

Can we soften some of their sharp edges?

Binding two factors via `fbind()`:

```{r}
library(foofactors)
a <- factor(c("character", "hits", "your", "eyeballs"))
b <- factor(c("but", "integer", "where it", "counts"))
```

Simply catenating two factors leads to a result that most don't expect.

```{r}
c(a, b)
```

The `fbind()` function glues two factors together and returns factor.

```{r}
fbind(a, b)
```

Often we want a table of frequencies for the levels of a factor. The base `table()` function returns an object of class `table`, which can be inconvenient for downstream work.  Processing with `as.data.frame()` can be helpful but it's a bit clunky.

```{r}
set.seed(1234)
x <- factor(sample(letters[1:5], size = 100, replace = TRUE))
table(x)
as.data.frame(table(x))
```

The `freq_out()` function returns a frequency table as a well-named `tbl_df`:

```{r}
freq_out(x)
```


### Reorder

Reorder is a function to change the order of the levels in a factor variable. The method reorders the levels based on the values of the second variable, usually numeric.

The `reorder` function returns the "scores" attribute which is the result of the function applied to the values of the second variable.

```{r}
test_factors <- ordered(c("a","a","b","b","c","c","a","b","c","a","c","a"))
test_vals <- seq(1,12)
output_reorder <- reorder(test_factors, test_vals, mean)
```

The output of the function has the entries in the atomic vector in the same order as given in the input but the levels are sorted in descending order according to the value of scores.

```{r}
attributes(output_reorder)$scores
levels(test_factors)
levels(output_reorder)
```

